---
title-block-banner: true
title: "EMA State Self Compassion Project"
description: "Test of the bipolar continuum hypothesis"
author: "Corrado Caudek (corrado.caudek@unifi.it)"
bibliography: bibliography.bib
csl: apa.csl
format: 
  html:
    code-line-numbers: true
    anchor-sections: true
    smooth-scroll: true
    grid:
      margin-width: 600px
editor: source
toc: True
date: last-modified
date-format: iso
---

# Introduction

The construct of self-compassion, a central concept in contemporary psychology, has been the subject of much scholarly debate. Traditionally, researchers have grappled with whether self-compassion is best understood as a unitary construct, represented by a bipolar continuum ranging from uncompassionate self-responding (UCS) to compassionate self-responding (CS), or as a configuration of two partly independent components. This debate is vital as it shapes our understanding of self-compassion and its role in psychological well-being. 

Building on this discussion, recent advancements in psychological research have emphasized the dynamic nature of psychological constructs, suggesting that they are not static or solely trait-based, but rather can be influenced by situational factors. This perspective is critical in the study of state self-compassion, an evolution of the original construct of self-compassion, which encompasses moment-to-moment variations in self-compassionate feelings and attitudes. The development of scales to measure state self-compassion (e.g., Raes et al., 2011) provides a tool to explore this dynamism.

In our study, we aim to contribute to the ongoing discourse by examining how situational factors, such as negative affect, influence the UCS and CS components of state self-compassion. This approach is inspired by the shift in emotion regulation research, where the focus has moved from the frequency of using strategies to the context-dependent effectiveness of these strategies (Gross, 1998; Aldao et al., 2015). By investigating how different situational factors impact UCS and CS, we seek to provide empirical evidence to the theoretical debate on the nature of the self-compassion construct.

Our hypothesis is twofold: If situational factors affect UCS and CS in parallel but opposing ways (i.e., an increase in UCS corresponds with a decrease in CS), this would lend support to the bipolar continuum model of self-compassion, as proposed by Neff (2003). This model posits that self-compassion is a singular, unified construct with varying degrees of expression. Alternatively, if situational factors differentially influence UCS and CS (for instance, if a factor impacts CS but leaves UCS unaffected, or vice versa), this would suggest that self-compassion comprises two distinct components, aligning with the perspectives of researchers like Muris et al. (2016), who argue for the duality of the construct.

This exploration is crucial for a deeper understanding of self-compassion, as it not only contributes to theoretical knowledge but also has practical implications. Understanding the situational dynamics of self-compassion can enhance psychological interventions and promote well-being by identifying specific contexts in which self-compassionate or uncompassionate responses are more likely to be elicited. Such insights could be pivotal in developing more nuanced and effective therapeutic strategies that cater to individual differences and situational variables, ultimately fostering a more compassionate self-relationship in diverse life situations.


## Decentering

Il costrutto di "decentering" nella psicologia si riferisce alla capacità di osservare i propri pensieri ed emozioni come eventi mentali temporanei e oggettivi, piuttosto che come riflessi accurati della realtà o aspetti immutabili del sé. Questo concetto è spesso associato alla mindfulness e alla terapia cognitivo-comportamentale, in particolare nelle pratiche di mindfulness e meditazione.

La relazione tra decentering e self-compassion è complessa e interconnessa. Entrambi i costrutti contribuiscono al benessere psicologico, ma lo fanno attraverso meccanismi leggermente diversi:

1. **Riduzione dell'Identificazione con Pensieri Negativi**: Il decentering aiuta a ridurre l'identificazione con pensieri e sentimenti negativi, permettendo agli individui di osservarli senza giudicarli o reagire in modo automatico. Questo distacco può creare uno spazio in cui la self-compassion può emergere, poiché gli individui possono riconoscere le proprie esperienze difficili con una maggiore gentilezza e comprensione.

2. **Gestione delle Emozioni Difficili**: La self-compassion, a sua volta, può facilitare il processo di decentering offrendo un approccio più gentile e accogliente verso le emozioni e i pensieri difficili. Quando gli individui sono in grado di trattare se stessi con compassione, possono trovare più facile osservare le proprie esperienze interne senza esserne sopraffatti.

3. **Risposta a Fallimenti e Critiche**: La self-compassion consente di affrontare fallimenti, errori e autocritiche con un atteggiamento di gentilezza e comprensione, piuttosto che con giudizio o rimprovero. Questo approccio può facilitare il decentering, poiché gli individui apprendono a vedere tali esperienze come comuni a tutti gli esseri umani, piuttosto che come riflessioni del proprio valore personale.

4. **Rafforzamento della Resilienza**: Entrambi contribuiscono alla resilienza psicologica. Il decentering aiuta a mantenere una prospettiva più equilibrata e obiettiva, mentre la self-compassion offre supporto emotivo e una risposta più positiva alle sfide.

5. **Mindfulness come Collegamento**: La mindfulness è un fattore chiave che collega decentering e self-compassion. La pratica della mindfulness incoraggia sia il decentering sia la self-compassion, promuovendo un'osservazione non giudicante delle esperienze interne e una maggiore gentilezza verso se stessi.

In sintesi, il decentering e la self-compassion sono entrambi importanti per il benessere psicologico e spesso si rafforzano a vicenda. La capacità di osservare i propri pensieri e sentimenti da una distanza (decentering) può facilitare un approccio più compassionevole verso se stessi, e viceversa, la self-compassion può rendere più facile l'adozione di una prospettiva decentered.


```{r}
#| include = FALSE
#| 
suppressPackageStartupMessages({
  library("here")
  library("targets")
  library("tidyverse")
  library("mice")
  library("lme4")
  library("brms")
  library("bayesplot")
  library("effectsize")
  library("scales")
  library("sjstats")
  library("sjPlot")
  library("sjmisc")
  library("emmeans")
})
```

## State Self-Compassion


## Method

### Task Design

### Compliance

```{r}
tar_load(compliance_results)
compliance_results
```

## Multilevel Reilabilities

```{r}
tar_load(reliabilities_sem)
```

### Positive State Self-Compassion

```{r}
reliabilities_sem[[1]]
```

### Negative State Self-Compassion

```{r}
reliabilities_sem[[2]]
```


## Negative Affect Variation Before and After the Exam

First exam.

```{r}
tar_load(list_params_first_exam_neg_aff_difference)
list_params_first_exam_neg_aff_difference
```

Second exam.

```{r}
tar_load(list_params_second_exam_neg_aff_difference)
list_params_second_exam_neg_aff_difference
```

## Comparison Negative Affect Pre-exam / Post-exam and the Other Days

Comparison between the negative affect either pre-exam and post-exam and the average negative affect in the other days.

Comparison other days and pre-exam.

```{r}
tar_load(list_params_no_exam_pre_neg_aff_difference)
list_params_no_exam_pre_neg_aff_difference
```

Comparison other days and post-exam.

```{r}
tar_load(list_params_no_exam_post_neg_aff_difference)
list_params_no_exam_post_neg_aff_difference
```


### Interpretation

In our multilevel analysis of State Self-Compassion, we observed significant 
reliability indices for both the positive and negative components. Focusing 
first on the positive aspect, the within-subject reliability 
($\tilde{\omega}^w$) was recorded at 0.626. This figure indicates a 
moderate level of consistency in individual responses across various 
measurement instances. In contrast, the between-subject reliability 
($\tilde{\omega}^b$), crucial for differentiating between individuals, was 
found to be 0.820. This higher value underscores the measure's effectiveness 
in capturing stable individual differences in self-compassion.
Moreover, the composite reliability of the overall positive score 
($\omega^{2L}$), which integrates both within and between subject 
variabilities, was calculated to be 0.79. This aligns with the framework 
proposed by Lai (2021) and confirms the measure's reliability in a 
comprehensive manner.
Turning to the negative component of State Self-Compassion, the 
within-subject reliability ($\tilde{\omega}^w$) was slightly higher at 
0.68, while the between-subject reliability ($\tilde{\omega}^b$) further 
increased to 0.88. The composite reliability for the overall negative 
component score ($\omega^{2L}$) was determined to be 0.83. These values, 
particularly the between-subject reliabilities exceeding the anticipated 
benchmark of approximately 0.8, emphasize the robustness of the measure.
The lower within-subject reliabilities for both components, while perhaps 
initially surprising, are understandable upon closer examination. These 
values reflect not only measurement error but also the inherent variability 
of State Self-Compassion as a dynamic construct, influenced by changing 
circumstances and internal states. Therefore, the lower within-subject 
reliabilities underscore the measure's sensitivity to these temporal changes 
within individuals.
In summary, the multilevel reliabilities of the State Self-Compassion measure, 
with robust between-subject reliability and lower, yet insightful, 
within-subject reliability, demonstrate its overall soundness and 
applicability in the present study. The composite reliability figures further 
confirm the measure's effectiveness in comprehensively capturing the 
construct of State Self-Compassion across different levels of analysis.


## Models' Comparisons

### Negative Component of State Self-Compassion

The final model for predicting the negative component of State Self-Compassion is as follows:

$$
\begin{equation}
\begin{aligned}
znsc &\sim \text{Student-t}(\mu, \sigma, \nu) \\
\mu &= \beta_{0} + \beta_{\text{dec\_moment}} \times \text{dec\_moment} + \beta_{\text{dec\_day}} \times \text{dec\_day} + \beta_{\text{dec\_person}} \times \text{dec\_person} + \\
&\quad \beta_{\text{na\_moment}} \times \text{na\_moment} + \beta_{\text{na\_day}} \times \text{na\_day} + \beta_{\text{na\_person}} \times \text{na\_person} + \\
&\quad \beta_{\text{context\_moment}} \times \text{context\_moment} + \beta_{\text{context\_day}} \times \text{context\_day} + \beta_{\text{context\_person}} \times \text{context\_person} + \\
&\quad b_{\text{user\_id}}[j] + b_{\text{bysubj\_day}}[k] \\
b_{\text{user\_id}}[j] &\sim \mathcal{N}(0, \Sigma_{\text{user\_id}}) \\
b_{\text{bysubj\_day}}[k] &\sim \mathcal{N}(0, \sigma^2_{\text{bysubj\_day}}) \\
\beta &\sim \text{priors1} \\
\sigma &\sim \text{Half-Cauchy}(0, \text{scale}) \\
\nu &\sim \text{Exponential}(\text{rate}) \\
\\
\text{Where:} & \\
\Sigma_{\text{user\_id}} &= \text{Full covariance matrix for random effects within user\_id} \\
\sigma^2_{\text{bysubj\_day}} &= \text{Var}(b_{\text{bysubj\_day}}[k])
\end{aligned}
\end{equation}
$$

### Results

We employed a Bayesian multilevel model to predict the negative component of the State Self-Compassion scale, which was measured multiple times for each participant across various occasions. The model included momentary negative affect (na), the evaluation of momentary context valence (context), and momentary decentering (dec) as predictors. Each of these predictors was decomposed into three components: within-person, between-day, and between-person.

#### Model Summary

The model was estimated using a Student's t-distribution with identity links for the mean (mu), scale (sigma), and degrees of freedom (nu). The analysis was based on 6741 observations, with the posterior distribution derived from 2000 post-warmup draws across two chains.

#### Group-Level Effects

- **By-Subject-Day Variability**: The standard deviation of the random intercepts for different days (`bysubj_day`) was estimated to be 0.02, suggesting minimal variability in the negative component of State Self-Compassion across days within subjects.
- **By-User Variability**: The model revealed substantial between-person variability in the intercept (0.50), indicating differences in baseline levels of the negative component of State Self-Compassion among participants. The variability in the within-person components of negative affect (`na_moment`: 0.06, `na_day`: 0.15), context (`context_moment`: 0.02, `context_day`: 0.09), and decentering (`dec_moment`: 0.08, `dec_day`: 0.14) suggests individual differences in how these factors relate to self-compassion on a moment-to-moment and day-to-day basis.

#### Population-Level Effects

- **Decentering**: Momentary (β = -0.12), daily (β = -0.18), and person-level (β = -0.37) decentering showed negative associations with the negative component of State Self-Compassion, indicating that higher levels of decentering are associated with lower levels of self-criticism or negative self-judgment.
- **Negative Affect**: Momentary (β = 0.13) and day-level (β = 0.17) negative affect were positively associated with the negative component, while person-level negative affect (β = 0.34) showed a stronger positive relationship. This suggests that both transient and enduring negative emotions contribute to increased self-criticism.
- **Context Valence**: Momentary context valence showed a weak negative association (β = -0.02), while daily (β = 0.01) and person-level context valence (β = 0.04) had minimal or slightly positive associations, indicating that the immediate context's perceived negativity might slightly increase self-critical thoughts.

#### Family Specific Parameters

- The estimated scale parameter (σ = 0.35) and degrees of freedom (ν = 4.20) of the Student's t-distribution reflect the variability and the heaviness of the tails in the model residuals, respectively.

#### Model Diagnostics

The model diagnostics indicate satisfactory convergence with Rhat values close to 1 for all parameters. The Bulk_ESS and Tail_ESS values suggest adequate effective sample sizes for reliable estimation and inference.

#### Output of brm()

`b_dec_moment`

```{r}
tar_load(stats_coef_dec_moment)
stats_coef_dec_moment
```

`b_dec_day`

```{r}
tar_load(stats_coef_dec_moment)
stats_coef_dec_day
```

`b_dec_person`

```{r}
tar_load(stats_coef_dec_moment)
stats_coef_dec_person
```


### Random effects

cor(dec_moment,dec_day) = 0.46: This is the correlation between the random effects for decentering at the moment level and the day level within the same individual. A positive correlation of 0.46 suggests that there is a moderate relationship between the variability in decentering across different days and within a single day for the same participant. In simpler terms, if a participant shows considerable variation in their decentering scores from day to day, they are also likely to show noticeable variation in decentering at different moments within the same day.

cor(na_moment,na_day) = 0.52: This coefficient represents the correlation between the random effects for negative affect at the moment level and the day level within the same individual. Similarly to the previous coefficient, a positive correlation of 0.52 indicates a moderate to strong relationship. It means that participants who tend to vary a lot in their negative affect scores between days are also likely to exhibit significant variation within a single day.

These correlations are important in understanding the dynamics of the psychological constructs you are studying (decentering and negative affect) in relation to time (both across days and within a single day). It suggests that there is a consistency in the way individuals vary in these constructs over time, both on a daily and momentary basis. 



