---
title-block-banner: true
title: "EMA State Self Compassion Project"
description: "Test of the bipolar continuum hypothesis"
author: "Corrado Caudek (corrado.caudek@unifi.it)"
bibliography: bibliography.bib
csl: apa.csl
format: 
  html:
    code-line-numbers: true
    anchor-sections: true
    smooth-scroll: true
    grid:
      margin-width: 600px
editor: source
toc: True
date: last-modified
date-format: iso
---

# Introduction

```{r}
#| include = FALSE
#| 
suppressPackageStartupMessages({
  library("here")
  library("targets")
  library("tidyverse")
  library("mice")
  library("lme4")
  library("brms")
  library("bayesplot")
  library("effectsize")
  library("scales")
  library("sjstats")
  library("sjPlot")
  library("sjmisc")
  library("emmeans")
})
```

## State Self-Compassion


## Method

### Task Design



## Multilevel Reilabilities

```{r}
tar_load(reliabilities_sem)
```

### Positive State Self-Compassion

```{r}
reliabilities_sem[[1]]
```

### Negative State Self-Compassion

```{r}
reliabilities_sem[[2]]
```

### Interpretation


In our multilevel analysis of State Self-Compassion, we observed significant 
reliability indices for both the positive and negative components. Focusing 
first on the positive aspect, the within-subject reliability 
($\tilde{\omega}^w$) was recorded at 0.626. This figure indicates a 
moderate level of consistency in individual responses across various 
measurement instances. In contrast, the between-subject reliability 
($\tilde{\omega}^b$), crucial for differentiating between individuals, was 
found to be 0.820. This higher value underscores the measure's effectiveness 
in capturing stable individual differences in self-compassion.
Moreover, the composite reliability of the overall positive score 
($\omega^{2L}$), which integrates both within and between subject 
variabilities, was calculated to be 0.79. This aligns with the framework 
proposed by Lai (2021) and confirms the measure's reliability in a 
comprehensive manner.
Turning to the negative component of State Self-Compassion, the 
within-subject reliability ($\tilde{\omega}^w$) was slightly higher at 
0.68, while the between-subject reliability ($\tilde{\omega}^b$) further 
increased to 0.88. The composite reliability for the overall negative 
component score ($\omega^{2L}$) was determined to be 0.83. These values, 
particularly the between-subject reliabilities exceeding the anticipated 
benchmark of approximately 0.8, emphasize the robustness of the measure.
The lower within-subject reliabilities for both components, while perhaps 
initially surprising, are understandable upon closer examination. These 
values reflect not only measurement error but also the inherent variability 
of State Self-Compassion as a dynamic construct, influenced by changing 
circumstances and internal states. Therefore, the lower within-subject 
reliabilities underscore the measure's sensitivity to these temporal changes 
within individuals.
In summary, the multilevel reliabilities of the State Self-Compassion measure, 
with robust between-subject reliability and lower, yet insightful, 
within-subject reliability, demonstrate its overall soundness and 
applicability in the present study. The composite reliability figures further 
confirm the measure's effectiveness in comprehensively capturing the 
construct of State Self-Compassion across different levels of analysis.

