---
title-block-banner: true
title: "EMA State Self Compassion Project"
description: "Test of the bipolar continuum hypothesis"
author: "Corrado Caudek (corrado.caudek@unifi.it)"
bibliography: bibliography.bib
csl: apa.csl
format: 
  html:
    code-line-numbers: true
    anchor-sections: true
    smooth-scroll: true
    grid:
      margin-width: 600px
editor: source
toc: True
date: last-modified
date-format: iso
---

# Introduction

The construct of self-compassion, a central concept in contemporary psychology, has been the subject of much scholarly debate. Traditionally, researchers have grappled with whether self-compassion is best understood as a unitary construct, represented by a bipolar continuum ranging from uncompassionate self-responding (UCS) to compassionate self-responding (CS), or as a configuration of two partly independent components. This debate is vital as it shapes our understanding of self-compassion and its role in psychological well-being. 

Building on this discussion, recent advancements in psychological research have emphasized the dynamic nature of psychological constructs, suggesting that they are not static or solely trait-based, but rather can be influenced by situational factors. This perspective is critical in the study of state self-compassion, an evolution of the original construct of self-compassion, which encompasses moment-to-moment variations in self-compassionate feelings and attitudes. The development of scales to measure state self-compassion (e.g., Raes et al., 2011) provides a tool to explore this dynamism.

In our study, we aim to contribute to the ongoing discourse by examining how situational factors, such as negative affect, influence the UCS and CS components of state self-compassion. This approach is inspired by the shift in emotion regulation research, where the focus has moved from the frequency of using strategies to the context-dependent effectiveness of these strategies (Gross, 1998; Aldao et al., 2015). By investigating how different situational factors impact UCS and CS, we seek to provide empirical evidence to the theoretical debate on the nature of the self-compassion construct.

Our hypothesis is twofold: If situational factors affect UCS and CS in parallel but opposing ways (i.e., an increase in UCS corresponds with a decrease in CS), this would lend support to the bipolar continuum model of self-compassion, as proposed by Neff (2003). This model posits that self-compassion is a singular, unified construct with varying degrees of expression. Alternatively, if situational factors differentially influence UCS and CS (for instance, if a factor impacts CS but leaves UCS unaffected, or vice versa), this would suggest that self-compassion comprises two distinct components, aligning with the perspectives of researchers like Muris et al. (2016), who argue for the duality of the construct.

This exploration is crucial for a deeper understanding of self-compassion, as it not only contributes to theoretical knowledge but also has practical implications. Understanding the situational dynamics of self-compassion can enhance psychological interventions and promote well-being by identifying specific contexts in which self-compassionate or uncompassionate responses are more likely to be elicited. Such insights could be pivotal in developing more nuanced and effective therapeutic strategies that cater to individual differences and situational variables, ultimately fostering a more compassionate self-relationship in diverse life situations.


```{r}
#| include = FALSE
#| 
suppressPackageStartupMessages({
  library("here")
  library("targets")
  library("tidyverse")
  library("mice")
  library("lme4")
  library("brms")
  library("bayesplot")
  library("effectsize")
  library("scales")
  library("sjstats")
  library("sjPlot")
  library("sjmisc")
  library("emmeans")
})
```

## State Self-Compassion


## Method

### Task Design

### Compliance

```{r}
tar_load(compliance_results)
compliance_results
```

## Multilevel Reilabilities

```{r}
tar_load(reliabilities_sem)
```

### Positive State Self-Compassion

```{r}
reliabilities_sem[[1]]
```

### Negative State Self-Compassion

```{r}
reliabilities_sem[[2]]
```


## Negative Affect Variation Before and After the Exam

First exam.

```{r}
tar_load(list_params_first_exam_neg_aff_difference)
list_params_first_exam_neg_aff_difference
```

Second exam.

```{r}
tar_load(list_params_second_exam_neg_aff_difference)
list_params_second_exam_neg_aff_difference
```

## Comparison Negative Affect Pre-exam / Post-exam and the Other Days

Comparison between the negative affect either pre-exam and post-exam and the average negative affect in the other days.

Comparison other days and pre-exam.

```{r}
tar_load(list_params_no_exam_pre_neg_aff_difference)
list_params_no_exam_pre_neg_aff_difference
```

Comparison other days and post-exam.

```{r}
tar_load(list_params_no_exam_post_neg_aff_difference)
list_params_no_exam_post_neg_aff_difference
```


### Interpretation


In our multilevel analysis of State Self-Compassion, we observed significant 
reliability indices for both the positive and negative components. Focusing 
first on the positive aspect, the within-subject reliability 
($\tilde{\omega}^w$) was recorded at 0.626. This figure indicates a 
moderate level of consistency in individual responses across various 
measurement instances. In contrast, the between-subject reliability 
($\tilde{\omega}^b$), crucial for differentiating between individuals, was 
found to be 0.820. This higher value underscores the measure's effectiveness 
in capturing stable individual differences in self-compassion.
Moreover, the composite reliability of the overall positive score 
($\omega^{2L}$), which integrates both within and between subject 
variabilities, was calculated to be 0.79. This aligns with the framework 
proposed by Lai (2021) and confirms the measure's reliability in a 
comprehensive manner.
Turning to the negative component of State Self-Compassion, the 
within-subject reliability ($\tilde{\omega}^w$) was slightly higher at 
0.68, while the between-subject reliability ($\tilde{\omega}^b$) further 
increased to 0.88. The composite reliability for the overall negative 
component score ($\omega^{2L}$) was determined to be 0.83. These values, 
particularly the between-subject reliabilities exceeding the anticipated 
benchmark of approximately 0.8, emphasize the robustness of the measure.
The lower within-subject reliabilities for both components, while perhaps 
initially surprising, are understandable upon closer examination. These 
values reflect not only measurement error but also the inherent variability 
of State Self-Compassion as a dynamic construct, influenced by changing 
circumstances and internal states. Therefore, the lower within-subject 
reliabilities underscore the measure's sensitivity to these temporal changes 
within individuals.
In summary, the multilevel reliabilities of the State Self-Compassion measure, 
with robust between-subject reliability and lower, yet insightful, 
within-subject reliability, demonstrate its overall soundness and 
applicability in the present study. The composite reliability figures further 
confirm the measure's effectiveness in comprehensively capturing the 
construct of State Self-Compassion across different levels of analysis.

